### 2022-09-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@5dd3632ebad73129d6d692b071900fe366ba949e 🚀

\(^b\): joint-SDM-like approach. Note that the row for PCA also applies
to kernel/probabilistic PCA, which are variations on the more general

@@

The popularity of graph embedding techniques in machine learning is more
than the search for structural invariants: graphs are discrete objects,
and machine learning techniques tend to handle continuous data better.
Bringing a sparse graph into a continuous, dense vector space (Xu, 2020)
opens up a broader variety of predictive algorithms, notably of the sort
that are able to predict events as probabilities (Murphy, 2022).
Furthermore, the projection of the graph itself is a representation that
can be learned; Runghen et al. (2021), for example, used a neural
network to learn the embedding of a network in which not all
interactions were known, based on the nodes' metadata. This example has
many parallels in ecology (see fig.~\ref{fig:embedding} \textbf{C}), in
which node metadata can be given by phylogeny or functional traits.
Rather than directly predicting biological rules (see \emph{e.g.}
Pichler et al., 2020 for an overview), which may be confounded by the
sparse nature of graph data, learning embeddings works in the
low-dimensional space that maximizes information about the network
structure. This approach is further justified by the observation, for
example, that the macro-evolutionary history of a network is adequately
represented by some graph embeddings (Random dot product graphs (RDPG);
see Dalla Riva \& Stouffer, 2016). In a recent publication, Strydom et
al. (2022) have used an embedding (based on RDPG) to project a metaweb
of trophic interactions between European mammals, and transferred this
information to mammals of Canada, using the phylogenetic distance
between related clades to infer the values in the latent sub-space into
which the European metaweb was projected. By performing the RDPG step on
re-constructed values, this approach yields a probabilistic trophic
metaweb for mammals of Canada based on knowledge of European species,
despite a limited (\(\approx\) 5\%) taxonomic overlap.

>>>>

\(^b\): joint-SDM-like approach. Given the need to evaluate different
methods on a problem-specific basis, the fact that a lot of methods have
not been used on network problems is an opportunity for benchmarking and
method development. Note that the row for PCA also applies to
kernel/probabilistic PCA, which are variations on the more general

@@

The popularity of graph embedding techniques in machine learning is more
than the search for structural invariants: graphs are discrete objects,
and machine learning techniques tend to handle continuous data better.
Bringing a sparse graph into a continuous, dense vector space (Xu, 2020)
opens up a broader variety of predictive algorithms, notably of the sort
that are able to predict events as probabilities (Murphy, 2022).
Furthermore, the projection of the graph itself is a representation that
can be learned; Runghen et al. (2021), for example, used a neural
network to learn the embedding of a network in which not all
interactions were known, based on the nodes' metadata. This example has
many parallels in ecology (see fig.~\ref{fig:embedding} \textbf{C}), in
which node metadata can be represented by phylogeny, abundance, or
functional traits. Using phylogeny as a source of information assumes
(or strives to capture) the action of evolutionary processes on network
structure, which at least for food webs have been well documented
\textbf{TK REF}; similarly, the use of functional traits assumes that
interactions can be infered from the knowledge of trait-matching rules,
which is similarly well supported in the empirical literature \textbf{TK
REF}. Relating this information to an embedding rather than a list of
networks measures would allow to capture their effect on the more
fundamental aspects of network structure; conversely, the absence of a
phylogenetic or functional signal may suggest that evolutionary/trait
processes are not strong drivers of network structure, therefore opening
a new way to perform hypothesis testing.

Rather than directly predicting biological rules (see \emph{e.g.}
Pichler et al., 2020 for an overview), which may be confounded by the
sparse nature of graph data, learning embeddings works in the
low-dimensional space that maximizes information about the network
structure. This approach is further justified by the observation, for
example, that the macro-evolutionary history of a network is adequately
represented by some graph embeddings (Random dot product graphs (RDPG);
see Dalla Riva \& Stouffer, 2016). In a recent publication, Strydom et
al. (2022) have used an embedding (based on RDPG) to project a metaweb
of trophic interactions between European mammals, and transferred this
information to mammals of Canada, using the phylogenetic distance
between related clades to infer the values in the latent sub-space into
which the European metaweb was projected. By performing the RDPG step on
re-constructed values, this approach yields a probabilistic trophic
metaweb for mammals of Canada based on knowledge of European species,
despite a limited (\(\approx\) 5\%) taxonomic overlap.
